Viscous Transport in Eroding Porous Media

### 2019-09-30 docs/transport/erosion_transport.tex
NM

Once streamlines in an eroded geometry are computed, we characterize the
transport by computing and analyzing three different metrics: the
tortuosity, the anomalous dispersion, and the pore size distribution.
The local tortuosity of a streamline connecting the inlet to the outlet
is the streamline's length normalized by the distance between the inlet
and the outlet.  In porous media, depending on several factors including
the porosity, the local tortuosity can be greater than
1.5~\cite{kop-kat-tim1996, mat-kha-koz2008}, or even
2~\cite{dud-koz-mat2011}.  The tortuosity of a geometry is defined by
averaging the local tortuosity over all streamlines that are initialized
at the inlet, and the geometry's tortuosity characterizes average
particle motions~\cite{hak-com-den2019}.  To characterize spreading, the
fluid dispersion is computed by calculating the variance of the
streamline lengths. In porous media, this spreading is often
super-dispersive~\cite{kan-dea-nun-bij-blu-jua2014, cus-hu-den1995,
dea-leb-den-tar-bol-dav2013}. Accurately resolving the streamlines near
the grains is crucial since the anomalous dispersion is caused by
streamlines spending time in both the high and low velocity
regimes~\cite{ber-sch2001}.  Finally, since they are required to
quantify velocity distributions~\cite{ali-par-wei-bre2017,
dea-qua-bir-jua2018}, channelization~\cite{sie-ili-pri-riv-gua2019},
connectivity~\cite{knu-car2005, wes-blo-gra2001}, and to develop network
models~\cite{bry-kin-mel1993, bry-mel-cad1993, bij-blu2006}, we form the
pore size distribution throughout the erosion process.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
Our first contribution is extending the Barycentric quadrature rule of
Barnett et al.~\cite{bar-wu-vee2015} so that high-order accurate
derivatives of solutions of the Stokes equations can be computed.  Using
this new quadrature rule, we accurately compute erosion rates and stably
simulate erosion.  Moreover, we use the quadrature method to visualize
the vorticity in the fluid bulk.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Limitations}
As in our previous work~\cite{qua-moo2018}, we consider two-dimensional
immobile eroding grains. However, this work is not limited by the
proximity of the grains. Compared to our previous work, we are able to
simulate much denser suspensions with a comparable number of
discretization points and without upsampling.  While we do investigate
transport in this work, we only consider passive particles that are
advected by the flow.  To model other physical processes, such as
contaminant transport, we plan to simulate a finite Peclet number
advection-diffusion equation in eroded geometries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Outline of the Paper}
In Section~\ref{sec:formulation}, we summarize the erosion model that we
describe in detail in our previous work~\cite{qua-moo2018}.  In
Section~\ref{sec:DLP}, we recast all the governing equations as layer
potentials defined in both $\RR^2$ and in $\CC$.
Section~\ref{sec:transport} describes measures for characterizing the
geometry, transport, and dispersion. Section~\ref{sec:method} describes
the numerical methods, with special attention paid to the new quadrature
method to compute the shear stress and vorticity.
Section~\ref{sec:results} presents numerical examples for a variety of
dense packings of bodies.  Finally, concluding remarks are made in
Section~\ref{sec:conclusions}.

>>>>

Once \nick{the streamlines are computed, we characterize
transport by analyzing} three different metrics: the tortuosity, the anomalous dispersion, and the pore size distribution. 
The local tortuosity of a streamline \nick{that connects the inlet to the outlet is defined as the streamline's length normalized by the linear inlet-to-outlet distance}.
In porous media, the local tortuosity can be greater than 1.5~\cite{kop-kat-tim1996, mat-kha-koz2008} or even 2~\cite{dud-koz-mat2011}, \nick{depending on several factors such as the porosity.} 
The tortuosity of a geometry is defined by averaging the local tortuosity over all streamlines initialized at the inlet, and the geometry's tortuosity characterizes average particle motions~\cite{hak-com-den2019}.  To characterize spreading, the fluid dispersion is computed by calculating the variance of the streamline lengths. In porous media, this spreading is often
super-dispersive~\cite{kan-dea-nun-bij-blu-jua2014, cus-hu-den1995,
dea-leb-den-tar-bol-dav2013}. \nick{Since anomalous dispersion results from streamlines spending time in both the high and low velocity}
regimes~\cite{ber-sch2001}, it is crucial to accurately resolve  streamlines near grain boundaries, as achieved in this work.
Finally, \nick{the distribution of pore sizes is required to} quantify velocity distributions~\cite{ali-par-wei-bre2017, dea-qua-bir-jua2018}, channelization~\cite{sie-ili-pri-riv-gua2019}, connectivity~\cite{knu-car2005, wes-blo-gra2001}, and to develop network models~\cite{bry-kin-mel1993, bry-mel-cad1993, bij-blu2006}. \nick{Hence, we construct the pore-size distribution throughout the erosion process.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
Our first contribution is extending the Barycentric quadrature rule of
Barnett et al.~\cite{bar-wu-vee2015} \nick{to accurately compute velocity gradients associated with Stokes flow.}
\nick{This new quadrature rule permits accurate and stable simulations of erosion, and, moreover, enables visualization of the vorticity field throughout.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Limitations}
As in our previous work~\cite{qua-moo2018}, we consider two-dimensional
immobile eroding grains. However, this work \nick{does not limit} the proximity of  grains. Compared to our previous work, we are able to
simulate much denser suspensions with a comparable number of
discretization points and without upsampling.  While we do investigate
transport in this work, we only consider passive particles that are
advected by the flow.  To model other physical processes, such as
contaminant transport, we plan to simulate a finite Peclet number
advection-diffusion equation in eroded geometries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Outline of the Paper}
In Section~\ref{sec:formulation}, we summarize the erosion model \nick{that is described in more detail in} previous work~\cite{qua-moo2018}.  In
Section~\ref{sec:DLP}, we recast all the governing equations as layer
potentials defined in both $\RR^2$ and in $\CC$.
Section~\ref{sec:transport} describes measures for characterizing the
geometry, transport, and dispersion. Section~\ref{sec:method} describes
the numerical methods, with special attention paid to the new quadrature
method \nick{for computing} shear stress and vorticity.
Section~\ref{sec:results} presents numerical examples for a variety of
dense packings of bodies.  Finally, concluding remarks are made in
Section~\ref{sec:conclusions}.
